---
title: "RAVE Pipeline - Surface Reconstruction & CT Co-Registration"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# This code block sets up the engine environment
# Please do not remove me
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
build_pipeline <- raveio::configure_knitr()
if(FALSE) {
  invisible(list2env(as.list(local({
    settings <- raveio::load_yaml(file.path(rstudioapi::getActiveProject(), "modules", 
                                     "surface_reconstruction", "settings.yaml"))
    print(names(settings))
    settings
  })), envir = environment()))
}
```

```{rave check_commandline_tools, language = "R", export = "cmd_tools", cue = "always"}

dry_run <- raveio::is_dry_run() || isTRUE(dryrun)

dcm2niix <- tryCatch({
  dcm2niix <- raveio::normalize_commandline_path(
    path = dcm2niix_path, 
    unset = raveio:::cmd_dcm2niix(error_on_missing = FALSE)
  )
  if(!isTRUE(file.exists(dcm2niix))) { dcm2niix <- NULL }
  dcm2niix
}, error = function(e){ NULL })

freesurfer <- tryCatch({
  freesurfer <- raveio::normalize_commandline_path(
    path = freesurfer_path, 
    unset = raveio:::cmd_freesurfer_home(error_on_missing = FALSE)
  )
  if(!isTRUE(dir.exists(freesurfer))) { freesurfer <- NULL }
  freesurfer
}, error = function(e){ NULL })


flirt <- tryCatch({
  fsl <- raveio::normalize_commandline_path(
    path = fsl_path,
    type = "fsl",
    unset = raveio:::cmd_fsl_home(error_on_missing = FALSE)
  )
  flirt <- NULL
  if(isTRUE(dir.exists(fsl))) {
    flirt <- file.path(fsl, "bin", "flirt")
  }
  flirt
}, error = function(e){ NULL })

cmd_tools <- list(
  dry_run = dry_run,
  dcm2niix = dcm2niix,
  freesurfer = freesurfer,
  flirt = flirt
)
```

```{rave load_subject, language = "R", export = "subject"}

subject <- raveio::RAVESubject$new(
  project_name = project_name, 
  subject_code = subject_code,
  strict = FALSE
)
print(subject)
```

```{rave check_subject_data, export = "check_result", cue = "always"}

msgs <- character(0L)
warns <- character(0L)

fs_path <- subject$freesurfer_path
if(!length(fs_path) || is.na(fs_path) || !isTRUE(dir.exists(fs_path))) {
  fs_path <- file.path(subject$path, "fs")
  fs_reconstructed <- FALSE
} else {
  fs_reconstructed <- threeBrain::check_freesurfer_path(
    fs_path,
    autoinstall_template = FALSE,
    check_volume = TRUE,
    check_surface = FALSE
  )
}


mri <- file.path(subject$preprocess_settings$raw_path, path_mri)
if(!skip_recon) {
  if(is.null(cmd_tools$dcm2niix) || is.null(cmd_tools$freesurfer)) {
    warns <- append(warns, "Cannot find command-line `dcm2niix` and/or `FreeSurfer`: the reconstruction will be skipped")
    skip_recon <- TRUE
  } else {
    if(!isTRUE(dir.exists(mri))) {
      warns <- append(warns, "No MRI folder found, the reconstruction will be skipped")
      skip_recon <- TRUE
    }
  }
  if(!skip_recon && fs_reconstructed) {
    warns <- append(warns, sprintf("Found existing FreeSurfer reconstructed directory; this directory will be renamed: %s", fs_path))
  }
}

ct <- file.path(subject$preprocess_settings$raw_path, path_ct)
if(!skip_coregistration) {
  if(is.null(cmd_tools$flirt)) {
    warns <- append(warns, "Cannot find FSL-flirt; the co-registration will be skipped")
    skip_coregistration <- TRUE
  } else if( (!fs_reconstructed && (is.null(cmd_tools$dcm2niix) || is.null(cmd_tools$freesurfer))) ) {
    warns <- append(warns, "Cannot find FreeSurfer reconstruction, nor at least one of the following commands exists: `dcm2niix`, `FreeSurfer`. The co-registration will be skipped anyway")
    skip_coregistration <- TRUE
  } else {
    if(!isTRUE(dir.exists(ct))) {
      warns <- append(warns, "No CT folder found, the co-registration will be skipped")
      skip_coregistration <- TRUE
    }
  }
}

if(!skip_recon) {
  msgs <- append(msgs, sprintf("New FreeSurfer reconstruction will be created from %s", mri))
  msgs <- append(msgs, sprintf("MRI default folder will be set: %s", path_mri))
}
if(!skip_coregistration) {
  msgs <- append(msgs, sprintf("CT will be co-registered to MRI for electrode localization; CT path: %s", ct))
  msgs <- append(msgs, sprintf("CT default folder will be set: %s", path_ct))
}

path_temp <- file.path(subject$preprocess_settings$raw_path, "rave-imaging")

path_log <- file.path(path_temp, "log")

check_result <- list(
  project_name = subject$project_name,
  subject_code = subject$subject_code,
  fs_path = fs_path,
  
  fs_reconstructed = fs_reconstructed,
  skip_recon = skip_recon,
  skip_coregistration = skip_coregistration,
  
  has_dcm2niix = !is.null(cmd_tools$dcm2niix),
  has_freesurfer = !is.null(cmd_tools$freesurfer),
  has_flirt = !is.null(cmd_tools$flirt),
  
  path_mri = mri,
  path_ct = ct,
  path_temp = path_temp,
  path_log = path_log,
  
  messages = msgs,
  warnings = warns
)


```

```{rave set_default_paths, export = "default_paths", cue = "always"}

if(isTRUE(dir.exists(check_result$path_mri))) {
  ravedash::logger("Setting default MRI folder: ", path_mri, level = "info")
  subject$set_default("raw_mri_path", path_mri, namespace = "surface_reconstruction")
}

if(isTRUE(dir.exists(check_result$path_ct))) {
  ravedash::logger("Setting default CT folder: ", path_ct, level = "info")
  subject$set_default("raw_ct_path", path_ct, namespace = "surface_reconstruction")
}

default_paths <- subject$get_default(c("raw_mri_path", "raw_ct_path"), namespace = "surface_reconstruction")
default_paths
```

```{rave convert_DICOM_to_Nifti_dryrun, export = "script_dcm2nii", cue = "always"}
require(ravedash)

dcm2niix <- as.character(cmd_tools$dcm2niix)
if(!isTRUE(file.exists(dcm2niix))) {
  dcm2niix <- "dcm2niix"
}

path_temp <- check_result$path_temp
indir_mri <- check_result$path_mri
indir_ct <- check_result$path_ct

outdir_mri <- file.path(path_temp, "inputs", "MRI")
outdir_ct <- file.path(path_temp, "inputs", "CT")


merge <- params$dcm2niix$merge %OF% c("Auto", "No", "Yes")
merge <- c("-m n ", "-m y ", "")[[which(c("No", "Yes", "Auto") == merge)]]

float <- params$dcm2niix$float %OF% c("Yes", "No")
float <- c('-p y ', '-p n ')[[which(c("Yes", "No") == float)]]

crop <- params$dcm2niix$crop %OF% c("No", "Yes", "Ignore")
crop <- c("-x n", "-x y", "-x i")[[which(c("No", "Yes", "Ignore") == crop)]]

args <- paste0(merge, float, crop)

path_log <- normalizePath(check_result$path_log, winslash = "/", mustWork = FALSE)
log_file <- strftime(Sys.time(), "log-import-mri-ct.log")

cmd_type <- "cmd"
cmd <- c(
  "#!/usr/bin/env sh",
  "",
  "# This is your dcm2niix binary path",
  sprintf("cmd_dcm2niix=%s", dcm2niix),
  "",
  "# Prepare log file",
  sprintf("log_dir=%s", shQuote(path_log)),
  sprintf("log_file=%s", shQuote(log_file)),
  'mkdir -p "$log_dir" && touch "$log_dir/$log_file" && echo "Started: $(date -u)" > "$log_dir/$log_file"',
  'echo --------------------------------------------------------',
  'echo Log file: "$log_dir/$log_file"',
  'echo --------------------------------------------------------'
)
error <- TRUE
if(length(indir_mri) == 1) {
  cmdadd <- c(
    "# ------------------- MRI ---------------------",
    "# Assign MRI input & output folders",
    sprintf("subj_mri_in=%s", shQuote(indir_mri, type = cmd_type)),
    sprintf("subj_mri_out=%s", shQuote(outdir_mri, type = cmd_type)),
    '[ -d "$subj_mri_out" ] && rm -r "$subj_mri_out"',
    'mkdir -p "$subj_mri_out"',
    ""
  )
  if(dir.exists(indir_mri)) {
    error <- FALSE
    cmdadd <- c(
      cmdadd,
      "# Convert MRI (DICOM => Nifti)",
      sprintf('$cmd_dcm2niix %s -o "$subj_mri_out" "$subj_mri_in" >> "$log_dir/$log_file" 2>&1', args)
    )
  } else if(grepl(pattern = "nii($|\\.gz$)", x = indir_mri, ignore.case = TRUE)){
    error <- FALSE
    cmdadd <- c(
      cmdadd,
      "# Copy MRI Nifti file to RAVE temporary directory",
      'cp -r "$subj_mri_in" "$subj_mri_out'
    )
  } else {
    cmdadd <- NULL
  }
  cmd <- c(cmd, cmdadd)
}

if(length(indir_ct) == 1) {
  cmdadd <- c(
    "",
    "# ------------------- CT ----------------------",
    "# Assign CT input & output folders",
    sprintf("subj_ct_in=%s", shQuote(indir_ct, type = cmd_type)),
    sprintf("subj_ct_out=%s", shQuote(outdir_ct, type = cmd_type)),
    '[ -d "$subj_ct_out" ] && rm -r "$subj_ct_out"',
    'mkdir -p "$subj_ct_out"',
    ""
  )
  if(dir.exists(indir_ct)) {
    error <- FALSE
    cmdadd <- c(
      cmdadd,
      "# Convert CT (DICOM => Nifti)",
      sprintf('$cmd_dcm2niix %s -o "$subj_ct_out" "$subj_ct_in" >> "$log_dir/$log_file" 2>&1', args)
    )
    cmd <- c(cmd, cmdadd)
  } else if(grepl(pattern = "nii($|\\.gz$)", x = indir_ct, ignore.case = TRUE)){
    error <- FALSE
    cmdadd <- c(
      cmdadd,
      "# Copy MRI Nifti file to RAVE temporary directory",
      'cp -r "$subj_ct_in" "$subj_ct_out"'
    )
  } else {
    cmdadd <- NULL
  }
  cmd <- c(cmd, cmdadd)
}

cmd <- c(cmd, 'echo Done. >> "$log_dir/$log_file" && echo Done.')

script_dcm2nii <- list(
  script = cmd,
  path_temp = path_temp,
  error = error,
  log_file = normalizePath(file.path(path_log, log_file), winslash = "/", mustWork = FALSE)
)

```

```{rave convert_DICOM_to_Nifti, export = "conversion_mri", cue = "always"}

if(!isFALSE(script_dcm2nii$error)) {
  stop("Cannot convert DICOM to Nifti, invalid MRI or CT files")
}

dry_run <- raveio::is_dry_run() || isTRUE(dryrun)

path_temp <- script_dcm2nii$path_temp
outdir_mri <- file.path(path_temp, "inputs", "MRI")
outdir_ct <- file.path(path_temp, "inputs", "CT")
script_path <- file.path(raveio::dir_create2(file.path(path_temp, "scripts")), "cmd-import-mri-ct.sh")

backup_dir <- raveio::dir_create2(file.path(path_temp, "scripts", "backups"))
backup_path <- raveio::backup_file(script_path, remove = TRUE)
if(!isFALSE(backup_path) && isTRUE(file.exists(backup_path))) {
  file.rename(backup_path, file.path(backup_dir, basename(backup_path)))
}

# save script to path
writeLines(script_dcm2nii$script, con = script_path)
cmd <- sprintf("bash %s", shQuote(normalizePath(script_path, mustWork = TRUE)))

if(!dry_run) {
  unlink(script_dcm2nii$log_file)
  system(cmd, wait = TRUE)
  raveio::backup_file(script_dcm2nii$log_file, remove = FALSE)
}

conversion_mri <- list(
  command = cmd,
  dry_run = dry_run,
  script_path = script_path
)
```

```{rave command_FreeSurfer_recon_dryrun, export = "script_recon", cue = "always"}
require(ravedash)

fs_home <- as.character(cmd_tools$freesurfer)

if(isTRUE(check_result$skip_recon)) {
  script_recon <- list(
    error = TRUE,
    reason = list(message = "FreeSurfer reconstruction is manually skipped")
  )
} else if(!isTRUE(dir.exists(fs_home))) {
  script_recon <- list(
    error = TRUE,
    reason = list(message = "Cannot find FreeSurfer home directory")
  )
} else {
  
  fs_home <- normalizePath(fs_home, winslash = "/")
  
  # FREESURFER_HOME=$R_CACHE/freesurfer
  # source $FREESURFER_HOME/SetUpFreeSurfer.sh
  
  path_temp <- check_result$path_temp
  fs_sdir <- normalizePath(file.path(check_result$fs_path, ".."), mustWork = FALSE)
  
  infile <- params$nii_t1
  infile <- file.path(path_temp, "inputs", "MRI", infile)
  if(length(infile) != 1 || is.na(infile) || !file.exists(infile)) {
    script_recon <- list(
      error = TRUE,
      reason = list(message = "No MRI input specified")
    )
  } else {
    
    infile <- normalizePath(infile, mustWork = TRUE, winslash = "/")
    
    autorecon_flags <- c(
      "-all", "-autorecon1", "-autorecon2", "-autorecon3",
      "-autorecon2-cp", "-autorecon2-wm", "-autorecon2-pial"
    )
    flag <- params$freesurfer$steps %OF% autorecon_flags
    
    path_log <- normalizePath(check_result$path_log, winslash = "/", mustWork = FALSE)
    log_file <- sprintf("log-fs-recon%s.log", flag)
    
    cmd_type <- "cmd"
    cmd <- c(
      "#!/usr/bin/env bash",
      "",
      "# Set FreeSurfer home directory & initialize",
      sprintf("FREESURFER_HOME=%s", shQuote(fs_home)),
      "source $FREESURFER_HOME/SetUpFreeSurfer.sh",
      "",
      sprintf("SUBJECTS_DIR=%s", shQuote(path_temp)),
      sprintf('mri_infile=%s', shQuote(infile)),
      # '[ -d "$SUBJECTS_DIR/fs" ] && rm -r "$SUBJECTS_DIR/fs"',
      'mkdir -p "$SUBJECTS_DIR"',
      "# Prepare log file",
      sprintf("log_dir=%s", shQuote(path_log)),
      sprintf("log_file=%s", shQuote(log_file)),
      'mkdir -p "$log_dir" && touch "$log_dir/$log_file" && echo "Started: $(date -u)" > "$log_dir/$log_file"',
      'echo --------------------------------------------------------',
      'echo Log file: "$log_dir/$log_file"',
      'echo --------------------------------------------------------',
      local({
        if(isTRUE( params$freesurfer$fresh_start)) {
          c(
            '# Force removing FreeSurfer path and re-run the reconstruction',
            '# You need original Nifti files to run',
            'if [ -d "$SUBJECTS_DIR/fs" ]; then',
            '  rm -r "$SUBJECTS_DIR/fs"',
            'fi',
            sprintf('recon-all -sd "$SUBJECTS_DIR" -sid fs -i "$mri_infile" %s >> "$log_dir/$log_file" 2>&1', flag)
          ) 
        } else {
          c(
            'if [ -d "$SUBJECTS_DIR/fs/mri" ]; then',
            '  # Use existing FreeSurfer directory to continue analysis',
            '  # You do not need original Nifti files to run',
            sprintf('  recon-all -sd "$SUBJECTS_DIR" -sid fs %s', flag),
            'else',
            '  # Remove invalid FreeSurfer path and re-run the reconstruction',
            '  # You need original Nifti files to run',
            '  if [ -d "$SUBJECTS_DIR/fs" ]; then',
            '    rm -r "$SUBJECTS_DIR/fs"',
            '  fi',
            sprintf('  recon-all -sd "$SUBJECTS_DIR" -sid fs -i "$mri_infile" %s >> "$log_dir/$log_file" 2>&1', flag),
            'fi'
          )
        }
      })
    )
    
    cmd <- c(cmd, 'echo Done. >> "$log_dir/$log_file" && echo Done.')
    script_recon <- list(
      script = cmd,
      path_temp = path_temp,
      flag = flag,
      error = FALSE,
      log_file = normalizePath(file.path(path_log, log_file), winslash = "/", mustWork = FALSE)
    )
  }
}

```

```{rave command_FreeSurfer_recon, export = "fs_recon", cue = "always"}

if(!isFALSE(script_recon$error)) {
  stop(c(script_recon$reason$message, "Cannot run FreeSurfer commands")[[1]])
}

dry_run <- raveio::is_dry_run() || isTRUE(dryrun)

path_temp <- script_recon$path_temp
script_path <- file.path(raveio::dir_create2(file.path(path_temp, "scripts")), 
                         sprintf("cmd-fs-recon%s.sh", script_recon$flag))

# Back up existing script
backup_dir <- raveio::dir_create2(file.path(path_temp, "scripts", "backups"))
backup_path <- raveio::backup_file(script_path, remove = TRUE)
if(!isFALSE(backup_path) && isTRUE(file.exists(backup_path))) {
  file.rename(backup_path, file.path(backup_dir, basename(backup_path)))
}

# save script to path
writeLines(script_recon$script, con = script_path)
cmd <- sprintf("bash %s", shQuote(normalizePath(script_path, mustWork = TRUE)))

if(!dry_run) {
  unlink(script_recon$log_file)
  system(cmd, wait = TRUE)
  raveio::backup_file(script_recon$log_file, remove = FALSE)
}

ravedash::logger("Setting default T1 nii file: ", params$nii_t1, level = "info")
subject$set_default("nii_t1", params$nii_t1, namespace = "surface_reconstruction")

fs_recon <- list(
  command = cmd,
  dry_run = dry_run,
  script_path = script_path
)
```

```{rave command_coreg_dryrun, export = "script_coreg", cue = "always"}
require(ravedash)

flirt <- as.character(cmd_tools$flirt)

if(isTRUE(file.exists(flirt))) {
  fsl_dir <- normalizePath(dirname(dirname(flirt)))
} else {
  fsl_dir <- "/usr/local/fsl"
  flirt <- "flirt"
}

nii_t1 <- file.path(check_result$path_temp, "inputs", "MRI", params$nii_t1)
nii_ct <- file.path(check_result$path_temp, "inputs", "CT", params$nii_ct)
outdir <- raveio::dir_create2(file.path(check_result$path_temp, "coregistration"))

error <- FALSE
err_msg <- NULL

if(!isTRUE(file.exists(nii_t1))) {
  error <- TRUE
  err_msg <- "Cannot find T1 file"
}
if (!isTRUE(file.exists(nii_ct))) {
  error <- TRUE
  err_msg <- c(err_msg, "Cannot find CT file")
}

nii_t1 <- normalizePath(nii_t1, winslash = "/", mustWork = FALSE)
nii_ct <- normalizePath(nii_ct, winslash = "/", mustWork = FALSE)

path_log <- normalizePath(check_result$path_log, winslash = "/", mustWork = FALSE)
log_file <- strftime(Sys.time(), "log-ct-coregistration-t1.log")

cmd <- c(
  "#!/usr/bin/env bash",
  "",
  "# FSL setup",
  sprintf('export FSLDIR=%s', shQuote(fsl_dir)),
  'source ${FSLDIR}/etc/fslconf/fsl.sh',
  'PATH=${FSLDIR}/bin:${PATH}',
  "",
  "# Set input T1 & CT files",
  sprintf("infile_ct=%s", shQuote(nii_ct)),
  sprintf("infile_t1=%s", shQuote(nii_t1)),
  sprintf("outdir=%s", shQuote(outdir)),
  "",
  "# Prepare log file",
  sprintf("log_dir=%s", shQuote(path_log)),
  sprintf("log_file=%s", shQuote(log_file)),
  'mkdir -p "$log_dir" && touch "$log_dir/$log_file" && echo "Started: $(date -u)" > "$log_dir/$log_file"',
  'echo --------------------------------------------------------',
  'echo Log file: "$log_dir/$log_file"',
  'echo --------------------------------------------------------',
  "",
  "# Run flirt",
  'flirt -in "$infile_ct" -ref "$infile_t1" \\',
  '  -out "$outdir/ct_in_t1.nii" -omat "$outdir/ct2t1.mat" \\',
  '  -interp trilinear -cost mutualinfo -dof 6 -searchcost mutualinfo \\',
  '  -searchrx -180 180 -searchry -180 180 -searchrz -180 180 -v >> "$log_dir/$log_file" 2>&1',
  'echo Done. >> "$log_dir/$log_file" && echo Done.'
)
    
script_coreg <- list(
  script = cmd,
  path_temp = check_result$path_temp,
  error = FALSE,
  reason = list(message = paste(err_msg, collapse = "; ")),
  log_file = normalizePath(file.path(path_log, log_file), winslash = "/", mustWork = FALSE)
)
```

```{rave command_coreg, export = "coreg_results", cue = "always"}

if(!isFALSE(script_coreg$error)) {
  stop(c(script_coreg$reason$message, "Cannot run flirt commands")[[1]])
}

dry_run <- raveio::is_dry_run() || isTRUE(dryrun)
if(check_result$skip_coregistration) {
  dry_run <- TRUE
}

path_temp <- script_coreg$path_temp
script_path <- file.path(raveio::dir_create2(file.path(path_temp, "scripts")), 
                         "cmd-ct2t1-flirt.sh")

# Back up existing script
backup_dir <- raveio::dir_create2(file.path(path_temp, "scripts", "backups"))
backup_path <- raveio::backup_file(script_path, remove = TRUE)
if(!isFALSE(backup_path) && isTRUE(file.exists(backup_path))) {
  file.rename(backup_path, file.path(backup_dir, basename(backup_path)))
}

# save script to path
writeLines(script_coreg$script, con = script_path)
cmd <- sprintf("bash %s", shQuote(normalizePath(script_path, mustWork = TRUE)))

if(!dry_run) {
  unlink(script_coreg$log_file)
  system(cmd, wait = TRUE)
  raveio::backup_file(script_coreg$log_file, remove = FALSE)
}

ravedash::logger("Setting default CT nii file: ", params$nii_ct, level = "info")
subject$set_default("nii_ct", params$nii_ct, namespace = "surface_reconstruction")

coreg_results <- list(
  command = cmd,
  dry_run = dry_run,
  script_path = script_path
)
```

## Build, Visualize, & Run

Please make sure the following code block is at the end of your pipeline file. This block will build the pipeline and generate a `make-surface_reconstruction.R` script with your pipeline markdown file. `RAVE` will use the generated pipeline script to execute the pipeline in the dashboard application, or in massive production mode.

```{r build, echo=FALSE, results='hide'}
build_pipeline(make_file = "make-surface_reconstruction.R")
```


Once the pipeline script `make-surface_reconstruction.R` is built, you can visualize and execute the pipeline without the need of re-knit this document. Notice we use `r` block instead of `rave`. (This is because the code blocks are not part of pipeline targets.)

```{r visualize, echo=FALSE}
Sys.setenv("RAVE_PIPELINE" = normalizePath("."))
raveio::pipeline_visualize()
```






