---
title: "RAVE Pipeline Markdown Template"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
# This code block sets up the engine environment
# Please do not remove me
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
build_pipeline <- raveio::configure_knitr()
if(FALSE){
  settings <- raveio::load_yaml(file.path(rstudioapi::getActiveProject(), "modules", "wavelet_module", "settings.yaml"))
  list2env(as.list(settings), envir = environment())
}

```

```{rave load_subject, language = "R", export = "subject", cue = "always"}

# check if project and subject are valid
stopifnot(grepl("^[a-zA-Z0-9_]{1,}$", project_name))
stopifnot(grepl("^[a-zA-Z0-9_]{1,}$", subject_code))

subject <- raveio::RAVESubject$new(
  project_name = project_name, 
  subject_code = subject_code
)
print(subject)
```

```{rave get_notch_filter_timestamp, language = "R", export = "notch_filtere_stamp", cue = "always"}
notch_filtere_stamp <- subject$get_default(namespace = "notch_filter", "parameters", default_if_missing = Sys.time())
```

```{rave check_prerequisite, language = "R", export = "notch_filtered_electrodes", cue = "always"}
electrodes <- subject$electrodes
etypes <- subject$electrode_types
notch_filtered_electrodes <- electrodes[
  subject$preprocess_settings$notch_filtered &
    etypes %in% c("LFP", "EKG", "Audio")
]
if(!length(notch_filtered_electrodes)) {
  stop("There is no electrode available to the wavelet module")
}
```

```{rave get_raw_sample_rates, language = "R", export = "sample_rates", cue = "always"}
sample_rates <- subject$preprocess_settings$sample_rates
electrodes <- subject$electrodes

sample_rates <- sapply(notch_filtered_electrodes, function(e){
  re <- sample_rates[electrodes == e]
  if(!length(re)) {
    stop("Electrode ", e, " does not have sample rate. The data might not be imported correctly and some configurations are missing.")
  }
  re[[1]]
})

```



```{rave check_kernel, language = "R", export = "kernels", cue = "always"}

freqs <- unlist(kernel_table$Frequency)
cycles <- unlist(kernel_table$Cycles)

if(any(is.na(freqs) | freqs <= 0)) {
  stop("The wavelet kernel table contains negative `Frequency` value(s)")
}

if(length(cycles) != 2) {
  if(length(freqs) != length(cycles)) {
    stop("The wavelet kernel table lengths are inconsistent: the lenth of `Cycles` must be either 2 (a range on which wavelet cycles that will be interpolated log-linearly), or the same length as `Frequency`.")
  }
}

if(any(is.na(cycles) | cycles <= 1)) {
  stop("The wavelet kernel table contains negative or 0 `Cycles` value(s): wavelet cycles must be greater equal than 1 (also integer numbers are strongly recommended)")
}

if(length(freqs) != 2 && length(cycles) == 2) {
  if(cycles[[1]] > cycles[[2]]) {
    stop("The wavelet kernel table: when `Cycles` is a range, (length of 2), it must be in non-decreasing order.")
  }
}

if(length(precision) != 1 || !precision %in% c("float", "double")) {
  precision <- "float"
}

kernels <- list(
  freqs = freqs,
  cycles = cycles,
  precision = precision
)

```


```{rave run_wavelet, language = "R", export = "wavelet_params", cue = "always"}
# If notch filter results not changed, we can skip the wavelet
force(notch_filtere_stamp)

blocks <- subject$preprocess_settings$blocks
electrodes <- notch_filtered_electrodes
srates <- sample_rates
compress_rates <- srates / target_sample_rate
kernels_precision <- kernels$precision
pre_decimate <- as.integer(pre_downsample)
if(pre_decimate < 1) {
  pre_decimate <- 1
} 

overall_progress <- dipsaus::progress2("Wavelet overall progress", max = 5, shiny_auto_close = TRUE)

# Create dir
overall_progress$inc("Creating directories")
subject$initialize_paths(include_freesurfer = FALSE)
raveio::dir_create2(file.path(subject$data_path, "power"))
raveio::dir_create2(file.path(subject$data_path, "phase"))
raveio::dir_create2(file.path(subject$data_path, "voltage"))

# prepare kernels
overall_progress$inc("Generating wavelet kernels")
ravetools <- asNamespace("ravetools")
generate_kernel <- ravetools[[sprintf("wavelet_kernels2_%s", kernels_precision)]]

sample_file = file.path(subject$preprocess_path, 'voltage', 
                     sprintf('electrode_%d.h5', electrodes[[1]]))


if(!file.exists(sample_file) || !raveio::h5_valid(sample_file)){
  stop("Electrode file is missing (preprocess, electrode ", electrodes[[1]], ")")
}
sample_names <- gsub("^/", "", raveio::h5_names(sample_file))

lapply(unique(srates), function(srate) {
  lapply(blocks, function(block){
    sample_name <- sprintf("notch/%s", block)
    if(!sample_name %in% sample_names) {
      stop(sprintf("I can find the imported signal file for Electrode %s, but cannot find any notch-filtered signal for block %s. The data file might be corrupted.", electrodes[[1]], block))
    }
    sample_data <- raveio::load_h5(sample_file, name = sample_name, ram = FALSE, read_only = TRUE)
    data_length <- length(sample_data)
    
    if(data_length <= 0) {
      stop(sprintf("Electrode %s has zero-length signal (/notch/%s). The data file might be corrupted.", electrodes[[1]], block))
    }
    
    if(pre_downsample > 1) {
      sample_data <- ravetools::decimate(
        sample_data[], pre_downsample, ftype = "fir")
      data_length <- length(sample_data)
    }
    
    generate_kernel(freqs = kernels$freqs, srate = srate / pre_decimate, wave_num = kernels$cycles, data_length = data_length)
  })
})



### Prepare for the wavelet file. Since users may specify different sample rate, we need to remove files if they exists
# 1. reference file
# ref_file = file.path(dirs$channel_dir, 'reference', sprintf('ref_%s.h5', dipsaus::deparse_svec(electrodes)))
# unlink(ref_file)

# 2. raw channel files and power/phase files
overall_progress$inc("Removing previously generated wavelet coefficients")
preproc <- raveio::RAVEPreprocessSettings$new(subject = subject$subject_id, read_only = TRUE)

for(e in electrodes) {
  preproc$data[[as.character(e)]]$has_wavelet <- FALSE
}
preproc$save()

data_root <- subject$data_path
lapply(electrodes, function(e){
  unlink(file.path(data_root, 'power', sprintf('%d.h5', e)))
  unlink(file.path(data_root, 'phase', sprintf('%d.h5', e)))
  unlink(file.path(data_root, 'voltage', sprintf('%d.h5', e)))
  for(block in blocks){
    unlink(file.path(data_root, 'cache', 'power', 'raw', block, sprintf('%d.fst', e)))
    unlink(file.path(data_root, 'cache', 'phase', 'raw', block, sprintf('%d.fst', e)))
  }
})



# load signals
overall_progress$inc("Applying wavelet (a.k.a. the long step)")
preprocess_dir <- subject$preprocess_path
dipsaus::lapply_async2(
  seq_along(electrodes), function(ii){
    
    e <- electrodes[[ii]]
    srate <- srates[[ii]]
    compress_rate <- compress_rates[[ii]]
    
    if(pre_decimate > 1) {
      compress_rate <- compress_rates[[ii]] / pre_decimate
      srate <- srate / pre_decimate
    }
    
    for(block in blocks){
      
      sorig <- raveio::load_h5(
        file = file.path(preprocess_dir, 'voltage', sprintf('electrode_%d.h5', e)),
        name = sprintf('/notch/%s', block),
        ram = TRUE
      )
      
      if(pre_decimate > 1) {
        s <- ravetools::decimate(sorig, pre_decimate, 
                                 ftype = "fir")
      } else {
        s <- sorig
      }
      data_length <- length(s)
      
      re <- ravetools::morlet_wavelet(
        data = s,
        freqs = kernels$freqs,
        srate = srate,
        wave_num = kernels$cycles,
        precision = kernels_precision
      )
      
      # Subset coefficients to save space
      ind <- floor(seq(1, data_length, by = compress_rate))
      
      if(kernels_precision == "float"){
        # Load all at once and subset is faster, but one signal is around 1-2GB, so...
        coef <- t(re[ind,,drop = FALSE])
        
        phase <- Arg(coef)
        power <- Mod(coef)^2
        
        re$.mode <- "readwrite"
        re$delete()
      } else {
        coef <- t(re$real[ind, , drop = FALSE] + 1i * re$imag[ind, , drop = FALSE])
        
        phase <- Arg(coef)
        power <- Mod(coef)^2
        re$real$.mode <- "readwrite"
        re$real$delete()
        re$imag$.mode <- "readwrite"
        re$imag$delete()
      }
      
      # Save power, phase, voltage
      fname <- sprintf('%d.h5', e)
      wavelet_h5chunk <- c(length(kernels$freqs), 128)
      
      # power
      raveio::save_h5(
        x = power,
        file = file.path(data_root, "power", fname),
        name = sprintf('/raw/power/%s', block),
        chunk = wavelet_h5chunk,
        replace = TRUE
      )
      # save_h5(
      #   x = 'noref',
      #   file = file.path(data_root, "power", fname),
      #   name = '/reference',
      #   chunk = 1,
      #   replace = TRUE, size = 1000
      # )
      
      # phase
      raveio::save_h5(
        x = phase,
        file = file.path(data_root, "phase", fname),
        name = sprintf('/raw/phase/%s', block),
        chunk = wavelet_h5chunk,
        replace = TRUE
      )
      
      # voltage
      raveio::save_h5(
        x = sorig,
        file = file.path(data_root, "voltage", fname),
        name = sprintf('/raw/voltage/%s', block),
        chunk = 1024,
        replace = TRUE
      )
    }
  },
  plan = FALSE,
  callback = function(ii){
    sprintf(sprintf("Applying wavelet|Electrode - %s", electrodes[[ii]]))
  }
)

overall_progress$inc("Saving configurations and update log files")

# reload preprocess settings in case some changes are not captured
preproc <- raveio::RAVEPreprocessSettings$new(subject = subject$subject_id, read_only = TRUE)

for(e in electrodes) {
  preproc$data[[as.character(e)]]$has_wavelet <- TRUE
}

wavelet_params <- list(
  channels = electrodes,
  electrodes = electrodes,
  downsample_to = target_sample_rate,
  target_srate = target_sample_rate,
  frequencies = kernels$freqs,
  wave_num = kernels$cycles,
  cycle = kernels$cycles,
  precision = kernels_precision,
  pre_downsample = pre_downsample
)

wavelet_logs <- as.list(preproc$data$wavelet_logs)
wavelet_logs[[length(wavelet_logs) + 1]] <- wavelet_params
preproc$data$wavelet_logs <- wavelet_logs

wavelet_params <- wavelet_params[c(
  "electrodes", "downsample_to", "frequencies",
  "cycle", "precision", "pre_downsample"
)]
wavelet_params$timestamp <- strftime(Sys.time(), usetz = TRUE)
preproc$data$wavelet_params <- wavelet_params
preproc$save()

subject$set_default(
  namespace = "wavelet_module",
  key = "parameters",
  wavelet_params
)

# generate reference table
raveio::safe_write_csv(
  file = file.path(subject$meta_path, "reference_noref.csv"),
  row.names = FALSE,
  data.frame(
    Electrode = subject$electrodes,
    Group = "Default",
    Reference = "noref",
    Type = "No Reference"
  )
)

# generate frequencies.csv
utils::write.csv(
  file = file.path(subject$meta_path, "frequencies.csv"),
  row.names = FALSE,
  data.frame(
    Frequency = kernels$freqs,
    Cycle = kernels$cycles,
    Method = "Wavelet"
  )
)

```


```{rave clear_cache, language = "R", export = "clear_cache", cue = "always"}
# clear all existing cache data in cache_dir

clear_cache <- raveio::clear_cached_files(
  subject_code = subject$subject_code
)

```

## Build, Visualize, & Run

Please make sure the following code block is at the end of your pipeline file. This block will build the pipeline and generate a `make-wavelet_module.R` script with your pipeline markdown file. `RAVE` will use the generated pipeline script to execute the pipeline in the dashboard application, or in massive production mode.

```{r build, echo=FALSE, results='hide'}
build_pipeline(make_file = "make-wavelet_module.R")
```


Once the pipeline script `make-wavelet_module.R` is built, you can visualize and execute the pipeline without the need of re-knit this document. Notice we use `r` block instead of `rave`. (This is because the code blocks are not part of pipeline targets.)

```{r visualize, echo=FALSE}
Sys.setenv("RAVE_PIPELINE" = normalizePath("."))
raveio::pipeline_visualize()
```






