---
title: "RAVE Pipeline Markdown Template"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# This code block sets up the engine environment
# Please do not remove me
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
build_pipeline <- raveio::configure_knitr()
if(FALSE) {
  invisible(list2env(as.list(local({
    settings <- raveio::load_yaml(file.path(rstudioapi::getActiveProject(), "modules", 
                                     "electrode_localization", "settings.yaml"))
    print(names(settings))
    settings
  })), envir = environment()))
}
```


```{rave load_FreeSurfer_LUT, export = "fslut", cue = "never"}
fslut_path <- system.file("palettes", "datacube2", "FreeSurferColorLUT.json", package = "threeBrain")
cmap <- threeBrain::load_colormap(fslut_path)

fslut <- list(
  cmap = cmap,
  labels = sapply(cmap$map, '[[', "Label")
)
```

```{rave load_subject, language = "R", export = "subject"}
subject <- raveio::RAVESubject$new(
  project_name = project_name, 
  subject_code = subject_code,
  strict = FALSE
)
print(subject)
print(subject$freesurfer_path)
```


```{rave find_CT_Nifti_files, language = "R", export = "ct_candidates"}
fs_path <- subject$freesurfer_path

# nii/nii.gz files in fs_path/../coregistration, or fs_path/coregistration
if(length(fs_path) != 1 || is.na(fs_path) || !dir.exists(fs_path)) {
  stop("Cannot find surface/volume reconstruction folder.")
}

f1 <- list.files(file.path(fs_path, "coregistration"), pattern = "nii(?:\\.gz)?$", 
                 recursive = FALSE, ignore.case = TRUE, include.dirs = FALSE,
                 full.names = FALSE, all.files = FALSE)
f2 <- list.files(file.path(fs_path, "..", "coregistration"), pattern = "nii(?:\\.gz)?$", 
                 recursive = FALSE, ignore.case = TRUE, include.dirs = FALSE,
                 full.names = FALSE, all.files = FALSE)
files <- c(f1, f2)
files[duplicated(files)] <- sprintf("%s (2)", files[duplicated(files)])

ct_candidates <- structure(files, paths = list(
  f1 = f1, f2 = f2
))
```

The following code block only run once only when starting a new localization

```{rave check_localization_plan, export = "plan_list", cue = "always"}
require(dipsaus)
# check if number of electrodes is consistent with the plan
count <- dipsaus::fastmap2()
count$n <- 0
count$labels <- list()

plan_table <- lapply(localization_plan, function(item) {
  dim <- as.integer(dipsaus::parse_svec(item$dimension, unique = FALSE, sep = "[,x]"))
  dim <- dim[!is.na(dim)]
  if(!length(dim) || any(dim <= 0)) { return(NULL) }
  ne <- prod(dim)
  if(ne <= 0) { return(NULL) }
  label <- as.character(item$label)
  if(length(label) != 1 || is.na(label) || trimws(label) == '') {
    label <- "NoLabel"
  }
  if(!label %in% names(count$labels)) {
    count$labels[[label]] <- 0
  }
  type <- item$type
  if(!isTRUE(type %in% raveio::LOCATION_TYPES)) {
    type <- raveio::LOCATION_TYPES[[1]]
  }
  re <- data.frame(
    Electrode = count$n + seq_len(ne),
    Label = sprintf("%s%d", label, count$labels[[label]] + seq_len(ne)),
    LabelPrefix = label,
    Dimension = paste(dim, collapse = "x"),
    LocationType = type
  )
  count$n <- count$n + ne
  count$labels[[label]] <- count$labels[[label]] + ne
  re
})
plan_table <- dipsaus::drop_nulls(plan_table)
plan_table <- do.call('rbind', unname(plan_table))

electrodes <- sort(subject$preprocess_settings$electrodes)
if(length(electrodes) != length(plan_table$Electrode)) {
  stop(sprintf("The electrode plan table (n=%d) has inconsistent length with registered electrode length (n=%d).", length(electrodes), length(plan_table$Electrode)))
}
plan_table$Electrode <- electrodes
# Load existing electrodes_unsaved.csv or electrodes.csv, generate plan table
files <- file.path(subject$meta_path, c("electrodes_unsaved.csv", "electrodes.csv"))
files <- files[file.exists(files)]

try({
  if(length(files)) {
    files <- files[[1]]
    electrode_table <- raveio::safe_read_csv(files)
    tname1 <- c("Electrode", "Coord_x", "Coord_y", "Coord_z")
    tname2 <- c("Electrode", "Coord_x", "Coord_y", "Coord_z", "Radius", "SurfaceType", 
                "Hemisphere", "FSIndex", "FSLabel", "MNI305_x", "MNI305_y", "MNI305_z")
    if(all(tname1 %in% names(electrode_table))) {
      tname2 <- tname2[tname2 %in% names(electrode_table)]
      electrode_table <- electrode_table[, tname2]
      if(!setequal(as.integer(electrode_table$Electrode), as.integer(plan_table$Electrode)) ||
         nrow(electrode_table) != nrow(plan_table)) {
        stop("Existing electrode table has different row numbers to the plan table")
      }
      plan_table <- merge(electrode_table, plan_table, by = "Electrode")
    }
  }
}, silent = TRUE)

plan_table$Coord_x %?<-% 0
plan_table$Coord_y %?<-% 0
plan_table$Coord_z %?<-% 0
plan_table$Radius %?<-% 1
plan_table$SurfaceType %?<-% "pial"
plan_table$MNI305_x %?<-% 0
plan_table$MNI305_y %?<-% 0
plan_table$MNI305_z %?<-% 0
plan_table$VertexNumber <- -1
plan_table$FSIndex %?<-% 0
plan_table$FSLabel %?<-% "Unknown"
plan_table$SurfaceElectrode <- plan_table$LocationType %in% c("ECoG")
plan_table$SignalType <- subject$preprocess_settings$electrode_types

plan_list <- split(plan_table, plan_table$LabelPrefix)

# save to "electrodes_unsaved.csv"
utils::write.csv(
  plan_table,
  file = file.path(subject$meta_path, "electrodes_unsaved.csv"),
  row.names = FALSE
)
```


```{rave load_brain, language = "R", export = "brain", cue = "always"}
brain <- threeBrain::freesurfer_brain2(
  fs_subject_folder = subject$freesurfer_path,
  subject_name = subject$subject_code
)
```


```{rave Loading_brain_and_CT_if_exists, language = "R", export = "ct_in_t1"}
force(subject)

has_ct <- FALSE
ct_path <- character(0L)
if(length(path_ct_in_t1)) {
  if(startsWith(path_ct_in_t1, "{")) {
    s <- strsplit(path_ct_in_t1, "/|\\\\")[[1]]
    s[[1]] <- raveio::glue(s[[1]])
    s <- do.call(file.path, as.list(s))
    ct_path <- normalizePath(s, mustWork = TRUE)
  } else {
    ct_path <- normalizePath(path_ct_in_t1, mustWork = TRUE)
  }
  has_ct <- TRUE
}


if(has_ct) {
  subject$set_default("ct_path", path_ct_in_t1, namespace = "electrode_localization")
  ct_in_t1 <- threeBrain:::read_nii2(ct_path)
} else {
  subject$set_default("ct_path", NULL, namespace = "electrode_localization")
  ct_in_t1 <- NA
}
```

```{rave generate_indicator, language = "R", export = "ct_exists"}
ct_exists <- isTRUE(!is.null(ct_in_t1) && is.list(ct_in_t1))
```

```{rave generate_localization_viewer, language = "R", export = "viewer", cue = "always"}

control_presets <- c("localization")
controllers <- list()
controllers[["Highlight Box"]] <- FALSE
controllers[["Overlay Coronal"]] <- TRUE
controllers[["Overlay Axial"]] <- TRUE
controllers[["Overlay Sagittal"]] <- TRUE

if(!is.null(ct_in_t1) && is.list(ct_in_t1)) {
  ct <- ct_in_t1
  ct_shift <- ct$get_center_matrix()
  ct_qform <- ct$get_qform()
  matrix_world <- brain$Torig %*% solve(brain$Norig) %*% ct_qform %*% ct_shift
  threeBrain::add_voxel_cube(brain, "CT", ct$get_data(), size = ct$get_size(), matrix_world = matrix_world)
  key <- seq(0, max(ct$get_range()))
  cmap <- threeBrain::create_colormap(gtype = "volume", dtype = "continuous", 
            key = key, value = key, color = c("white", "green", "darkgreen"))
  controllers[["Left Opacity"]] <- 0.4
  controllers[["Right Opacity"]] <- 0.4
  controllers[["Voxel Type"]] <- "CT"
  controllers[["Voxel Display"]] <- "normal"
  controllers[["Voxel Min"]] <- 3000
  controllers[["Edit Mode"]] <- "CT/volume"
  
  viewer <- brain$plot(control_presets = control_presets, voxel_colormap = cmap, 
            controllers = controllers)
} else {
  controllers[["Edit Mode"]] <- "MRI slice"
        
  controllers[["Left Opacity"]] <- 0.1
  controllers[["Right Opacity"]] <- 0.1
  viewer <- brain$plot(control_presets = control_presets, controllers = controllers)
}

# If running in interactive mode
if(interactive()) {
  print(viewer)
}

```

```
# template_brain <- threeBrain::merge_brain(template_subject = "N27")
# atlas <- template_brain$template_object$atlases$aparc_aseg$group
# dim <- atlas$get_data("datacube_dim_Atlas - aparc_aseg (N27)")
# cube <- atlas$get_data("datacube_value_Atlas - aparc_aseg (N27)")
# dim(cube) <- dim
```

```{rave merge_localization_list, export = "localization_result_initial"}
require(dipsaus)
re <- lapply(localization_list, function(item) {
  
  item$FSIndex %?<-% 0
  item$FSLabel %?<-% "Unknown"
  item$Radius %?<-% 1
  
  tbl <- data.frame(
    Electrode = item$Electrode,
    Coord_x = item$Coord_x, Coord_y = item$Coord_y, Coord_z = item$Coord_z,
    Label = item$Label, LabelPrefix = item$LabelPrefix,
    Dimension = item$Dimension, 
    LocationType = item$LocationType, Radius = item$Radius, 
    MNI305_x = item$MNI305_x, MNI305_y = item$MNI305_y, MNI305_z = item$MNI305_z,
    FSIndex = item$FSIndex, FSLabel = item$FSLabel
  )
  if(!nrow(tbl)) { return(NULL) }
  tbl
})
re <- do.call("rbind", dipsaus::drop_nulls(re))
rownames(re) <- NULL
re <- re[order(re$Electrode), ]
re$SurfaceElectrode <- re$LocationType %in% c("ECoG")
re$SurfaceType <- "pial"
re$VertexNumber <- -1

# Calculate T1, MNI152
empty_sel <- (re$Coord_x)^2 + (re$Coord_y)^2 + (re$Coord_z)^2
empty_sel <- is.na(empty_sel) | empty_sel == 0
tkrRAS <- rbind(re$Coord_x, re$Coord_y, re$Coord_z, 1)
t1 <- brain$Norig %*% solve(brain$Torig) %*% tkrRAS
mni305 <- brain$xfm %*% brain$Norig %*% solve(brain$Torig) %*% tkrRAS
mni152 <- raveio:::MNI305_to_MNI152 %*% mni305

mni305[, empty_sel] <- 0
mni152[, empty_sel] <- 0
t1[, empty_sel] <- 0

re$MNI305_x <- mni305[1, ]
re$MNI305_y <- mni305[2, ]
re$MNI305_z <- mni305[3, ]
re$T1R <- t1[1, ]
re$T1A <- t1[2, ]
re$T1S <- t1[3, ]
re$MNI152_x <- mni152[1, ]
re$MNI152_y <- mni152[2, ]
re$MNI152_z <- mni152[3, ]

# save to electrodes_unsaved.csv
save_path <- file.path(subject$meta_path, "electrodes_unsaved.csv")
utils::write.csv(re, save_path, row.names = FALSE)

localization_result_initial <- re
```

Make sure you run target `localization_result_initial` first

```{rave calculate_std141_mappings, export = "localization_result_final"}
re <- utils::read.csv(file.path(subject$meta_path, "electrodes_unsaved.csv"))
brain$set_electrodes(electrodes = re)
re <- brain$calculate_template_coordinates(save_to = FALSE)

# check Hemisphere, make sure no NAs
sel <- is.na(re$Hemisphere) | !re$Hemisphere %in% c("left", "right")
if(any(sel)) {
  is_left <- grepl(pattern = "(left|lh-)", re$FSLabel, ignore.case = TRUE)
  is_right <- grepl(pattern = "(right|rh-)", re$FSLabel, ignore.case = TRUE)
  re$Hemisphere[sel & (is_left | re$MNI305_x < 0)] <- "left"
  re$Hemisphere[sel & !is_left & (is_right | re$MNI305_x > 0)] <- "right"
}

localization_result_final <- re

```



## Build, Visualize, & Run

Please make sure the following code block is at the end of your pipeline file. This block will build the pipeline and generate a `make-electrode_localization.R` script with your pipeline markdown file. `RAVE` will use the generated pipeline script to execute the pipeline in the dashboard application, or in massive production mode.

```{r build, echo=FALSE, results='hide'}
build_pipeline(make_file = "make-electrode_localization.R")
```


Once the pipeline script `make-electrode_localization.R` is built, you can visualize and execute the pipeline without the need of re-knit this document. Notice we use `r` block instead of `rave`. (This is because the code blocks are not part of pipeline targets.)

```{r visualize, echo=FALSE}
Sys.setenv("RAVE_PIPELINE" = normalizePath("."))
raveio::pipeline_visualize()
```







