---
title: "RAVE Pipeline Markdown Template"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# This code block sets up the engine environment
# Please do not remove me
raveio::pipeline_setup_rmd("electrode_localization")
```


```{rave load_FreeSurfer_LUT, export = "fslut", cue = "never"}
fslut_path <- system.file("palettes", "datacube2", "FreeSurferColorLUT.json", package = "threeBrain")
cmap <- threeBrain::load_colormap(fslut_path)

fslut <- list(
  cmap = cmap,
  labels = sapply(cmap$map, '[[', "Label")
)
```

```{rave load_subject, language = "R", export = "subject", format = "rave-subject"}
subject <- raveio::RAVESubject$new(
  project_name = project_name, 
  subject_code = subject_code,
  strict = FALSE
)
```


```{rave find_CT_Nifti_files, language = "R", export = "ct_candidates"}
fs_path <- subject$freesurfer_path

# nii/nii.gz files in fs_path/../coregistration, or fs_path/coregistration
if(length(fs_path) != 1 || is.na(fs_path) || !dir.exists(fs_path)) {
  stop("Cannot find surface/volume reconstruction folder.")
}

f1 <- list.files(file.path(fs_path, "coregistration"), pattern = "nii(?:\\.gz)?$", 
                 recursive = FALSE, ignore.case = TRUE, include.dirs = FALSE,
                 full.names = FALSE, all.files = FALSE)
f2 <- list.files(file.path(fs_path, "..", "coregistration"), pattern = "nii(?:\\.gz)?$", 
                 recursive = FALSE, ignore.case = TRUE, include.dirs = FALSE,
                 full.names = FALSE, all.files = FALSE)
files <- c(f1, f2)
files[duplicated(files)] <- sprintf("%s (2)", files[duplicated(files)])

ct_candidates <- structure(files, paths = list(
  f1 = f1, f2 = f2
))
```

The following code block only run once only when starting a new localization

```{rave check_localization_plan, export = "plan_list", cue = "always"}
require(dipsaus)
# check if number of electrodes is consistent with the plan
count <- dipsaus::fastmap2()
count$n <- 0
count$labels <- list()

plan_list <- NULL
plan_table <- lapply(localization_plan, function(item) {
  dim <- as.integer(dipsaus::parse_svec(item$dimension, unique = FALSE, sep = "[,x]"))
  dim <- dim[!is.na(dim)]
  if(!length(dim) || any(dim <= 0)) { return(NULL) }
  ne <- prod(dim)
  if(ne <= 0) { return(NULL) }
  label <- as.character(item$label)
  if(length(label) != 1 || is.na(label) || trimws(label) == '') {
    label <- "NoLabel"
  }
  if(!label %in% names(count$labels)) {
    count$labels[[label]] <- 0
  }
  type <- item$type
  if(!isTRUE(type %in% raveio::LOCATION_TYPES)) {
    type <- raveio::LOCATION_TYPES[[1]]
  }
  re <- data.frame(
    Electrode = count$n + seq_len(ne),
    Label = sprintf("%s%d", label, count$labels[[label]] + seq_len(ne)),
    LabelPrefix = label,
    Dimension = paste(dim, collapse = "x"),
    LocationType = type
  )
  count$n <- count$n + ne
  count$labels[[label]] <- count$labels[[label]] + ne
  re
})
plan_table <- dipsaus::drop_nulls(plan_table)

if(length(plan_table)) {
  plan_table <- do.call('rbind', unname(plan_table))
  
  electrodes <- sort(subject$preprocess_settings$electrodes)
  if(length(electrodes) == 0) {
    electrodes <- plan_table$Electrode
  }
  if(length(electrodes) != length(plan_table$Electrode)) {
    stop(sprintf("The electrode plan table (n=%d) has inconsistent length with registered electrode length (n=%d).", length(electrodes), length(plan_table$Electrode)))
  }
  plan_table$Electrode <- electrodes
  # Load existing electrodes_unsaved.csv or electrodes.csv, generate plan table
  files <- file.path(subject$meta_path, c("electrodes_unsaved.csv", "electrodes.csv"))
  files <- files[file.exists(files)]
  
  try({
    if(length(files)) {
      files <- files[[1]]
      electrode_table <- raveio::safe_read_csv(files)
      tname1 <- c("Electrode", "Coord_x", "Coord_y", "Coord_z")
      tname2 <- c("Electrode", "Coord_x", "Coord_y", "Coord_z", "Radius", "SurfaceType", 
                  "Hemisphere", "FSIndex", "FSLabel", "MNI305_x", "MNI305_y", "MNI305_z")
      if(all(tname1 %in% names(electrode_table))) {
        tname2 <- tname2[tname2 %in% names(electrode_table)]
        electrode_table <- electrode_table[, tname2]
        if(!setequal(as.integer(electrode_table$Electrode), as.integer(plan_table$Electrode)) ||
           nrow(electrode_table) != nrow(plan_table)) {
          stop("Existing electrode table has different row numbers to the plan table")
        }
        plan_table <- merge(electrode_table, plan_table, by = "Electrode")
      }
    }
  }, silent = TRUE)
  
  plan_table$Coord_x %?<-% 0
  plan_table$Coord_y %?<-% 0
  plan_table$Coord_z %?<-% 0
  plan_table$Radius %?<-% 1
  plan_table$SurfaceType %?<-% "pial"
  plan_table$MNI305_x %?<-% 0
  plan_table$MNI305_y %?<-% 0
  plan_table$MNI305_z %?<-% 0
  plan_table$VertexNumber <- -1
  plan_table$FSIndex %?<-% 0
  plan_table$FSLabel %?<-% "Unknown"
  plan_table$SurfaceElectrode <- plan_table$LocationType %in% c("ECoG")
  etypes <- subject$preprocess_settings$electrode_types
  if(!length(etypes)) {
    etypes <- "LFP"
  }
  plan_table$SignalType <- etypes
  
  plan_list <- split(plan_table, plan_table$LabelPrefix)
  
  # save to "electrodes_unsaved.csv"
  raveio::dir_create2(subject$meta_path)
  utils::write.csv(
    plan_table,
    file = file.path(subject$meta_path, "electrodes_unsaved.csv"),
    row.names = FALSE
  )
}


```


```{rave load_brain, language = "R", export = "brain", cue = "always"}
brain <- threeBrain::freesurfer_brain2(
  fs_subject_folder = subject$freesurfer_path,
  subject_name = subject$subject_code
)
```


```{rave Loading_brain_and_CT_if_exists, language = "R", export = "localize_data"}
force(subject)

resolve_path <- function(path) {
  if(length(path) != 1) {
    stop("Cannot resolve path: ", paste(format(path), collapse = "\n"))
  }
  if(startsWith(path, "{")) {
    s <- strsplit(path, "/|\\\\")[[1]]
    s[[1]] <- raveio::glue(s[[1]])
    s <- do.call(file.path, as.list(s))
    path <- normalizePath(s, mustWork = TRUE)
  } else {
    path <- normalizePath(path, mustWork = TRUE)
  }
  path
}

has_ct <- FALSE
ct_path <- character(0L)
mri_path <- character(0L)
mri_data <- NULL
transform_matrix <- NULL
if(length(path_ct)) {
  ct_path <- resolve_path(path_ct)
  has_ct <- TRUE
  subject$set_default("path_ct", path_ct, namespace = "electrode_localization")
  ct_data <- threeBrain:::read_nii2(ct_path)
  
  # check if method is MRI
  if(identical(tolower(transform_space), "fsl")) {
    transform_space <- "fsl"
    mri_path <- resolve_path(path_mri)
    subject$set_default("path_mri", path_mri, namespace = "electrode_localization")
    mri_data <- threeBrain:::read_nii2(mri_path, head_only = TRUE)
    
    transform_matrix <- as.matrix(read.table(resolve_path(path_transform)))
    dimnames(transform_matrix) <- NULL
    
    if(length(transform_matrix) != 16L || !is.numeric(transform_matrix)) {
      stop("Invalid transform matrix. Must be a 4x4 matrix.")
    }
    subject$set_default("path_transform", path_transform, namespace = "electrode_localization")
    
    
  } else {
    transform_space <- "resampled"
  }
  
} else {
  subject$set_default("path_ct", NULL, namespace = "electrode_localization")
  ct_data <- NULL
  transform_space <- "no_ct"
}

subject$set_default("transform_space", transform_space, namespace = "electrode_localization")

localize_data <- list(
  transform_space = transform_space,
  ct_data = ct_data,
  ct_path = ct_path,
  mri_path = mri_path,
  mri_data = mri_data,
  transform_matrix = transform_matrix
)
```

```{rave generate_indicator, language = "R", export = "ct_exists"}
ct_exists <- isTRUE(!is.null(localize_data$ct_data) && is.list(localize_data$ct_data))
```

```{rave generate_localization_viewer, language = "R", export = "viewer", cue = "always"}

force(ct_exists)
viewer <- brain$localize(
  ct_path = localize_data$ct_data, 
  transform_space = localize_data$transform_space,
  transform_matrix = localize_data$transform_matrix,
  mri_path = localize_data$mri_path
)

# If running in interactive mode
if(interactive()) {
  print(viewer)
}

```

```
# template_brain <- threeBrain::merge_brain(template_subject = "N27")
# atlas <- template_brain$template_object$atlases$aparc_aseg$group
# dim <- atlas$get_data("datacube_dim_Atlas - aparc_aseg (N27)")
# cube <- atlas$get_data("datacube_value_Atlas - aparc_aseg (N27)")
# dim(cube) <- dim
```

```{rave merge_localization_list, export = "localization_result_initial"}
localization_result_initial <- NULL
re <- lapply(localization_list, function(item) {
  
  item$FSIndex %?<-% 0
  item$FSLabel %?<-% "Unknown"
  item$Radius %?<-% 1
  
  tbl <- data.frame(
    Electrode = item$Electrode,
    Coord_x = item$Coord_x, Coord_y = item$Coord_y, Coord_z = item$Coord_z,
    Label = item$Label, LabelPrefix = item$LabelPrefix,
    Dimension = item$Dimension, 
    LocationType = item$LocationType, Radius = item$Radius, 
    MNI305_x = item$MNI305_x, MNI305_y = item$MNI305_y, MNI305_z = item$MNI305_z,
    FSIndex = item$FSIndex, FSLabel = item$FSLabel
  )
  if(!nrow(tbl)) { return(NULL) }
  tbl
})
re <- do.call("rbind", dipsaus::drop_nulls(re))

if(length(re) && nrow(re)) {
  rownames(re) <- NULL
  re <- re[order(re$Electrode), ]
  re$SurfaceElectrode <- re$LocationType %in% c("ECoG")
  re$SurfaceType <- "pial"
  re$VertexNumber <- -1
  
  # Calculate T1, MNI152
  empty_sel <- (re$Coord_x)^2 + (re$Coord_y)^2 + (re$Coord_z)^2
  empty_sel <- is.na(empty_sel) | empty_sel == 0
  tkrRAS <- rbind(re$Coord_x, re$Coord_y, re$Coord_z, 1)
  mr_voxel <- solve(brain$Torig) %*% tkrRAS
  t1 <- brain$Norig %*% mr_voxel
  mni305 <- brain$xfm %*% brain$Norig %*% solve(brain$Torig) %*% tkrRAS
  mni152 <- raveio:::MNI305_to_MNI152 %*% mni305
  
  mni305[, empty_sel] <- 0
  mni152[, empty_sel] <- 0
  t1[, empty_sel] <- 0
  mr_voxel[, empty_sel] <- 0
  
  re$MNI305_x <- mni305[1, ]
  re$MNI305_y <- mni305[2, ]
  re$MNI305_z <- mni305[3, ]
  re$T1R <- t1[1, ]
  re$T1A <- t1[2, ]
  re$T1S <- t1[3, ]
  re$MNI152_x <- mni152[1, ]
  re$MNI152_y <- mni152[2, ]
  re$MNI152_z <- mni152[3, ]
  re$MRVoxel_I <- round(mr_voxel[1, ])
  re$MRVoxel_J <- round(mr_voxel[2, ])
  re$MRVoxel_K <- round(mr_voxel[3, ])
  
  # save to electrodes_unsaved.csv
  save_path <- file.path(subject$meta_path, "electrodes_unsaved.csv")
  raveio::dir_create2(dirname(save_path))
  utils::write.csv(re, save_path, row.names = FALSE)
  
  localization_result_initial <- re
}
```

Make sure you run target `localization_result_initial` first

```{rave calculate_std141_mappings, export = "localization_result_final"}
src <- file.path(subject$meta_path, "electrodes_unsaved.csv")
if(file.exists(src)) {
  
  localization_result_final <- utils::read.csv(file.path(subject$meta_path, "electrodes_unsaved.csv"))
  
  try({
    re <- localization_result_final
    brain$set_electrodes(electrodes = re)
    re <- brain$calculate_template_coordinates(save_to = FALSE)
    
    # check Hemisphere, make sure no NAs
    sel <- is.na(re$Hemisphere) | !re$Hemisphere %in% c("left", "right")
    if(any(sel)) {
      is_left <- grepl(pattern = "(left|lh-)", re$FSLabel, ignore.case = TRUE)
      is_right <- grepl(pattern = "(right|rh-)", re$FSLabel, ignore.case = TRUE)
      re$Hemisphere[sel & (is_left | re$MNI305_x < 0)] <- "left"
      re$Hemisphere[sel & !is_left & (is_right | re$MNI305_x > 0)] <- "right"
    }
    
    localization_result_final <- re
  })
  
} else {
  localization_result_final <- NULL
}

```



## Build, Visualize, & Run

Please make sure the following code block is at the end of your pipeline file. This block will build the pipeline and generate a `make-electrode_localization.R` script with your pipeline markdown file. `RAVE` will use the generated pipeline script to execute the pipeline in the dashboard application, or in massive production mode.

```{r build, echo=FALSE, results='hide'}
build_pipeline(make_file = "make-electrode_localization.R")
```


Once the pipeline script `make-electrode_localization.R` is built, you can visualize and execute the pipeline without the need of re-knit this document. Notice we use `r` block instead of `rave`. (This is because the code blocks are not part of pipeline targets.)

```{r visualize, echo=FALSE}
Sys.setenv("RAVE_PIPELINE" = normalizePath("."))
raveio::pipeline_visualize()
```







